---
title: "Assignment 1 - Language Development in ASD - Power and simulations"
author: "[YOUR NAME]"
date: "[DATE]"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(simr)
```

## Welcome to the third exciting part of the Language Development in ASD exercise

In this part of the assignment, we try to figure out how a new study should be planned (i.e. how many participants?) in order to have enough power to replicate the findings (ensuring our sample size is adequate, our alpha at 0.05 and our beta at 0.8):
1- if we trust the estimates of the current study. Report the power analysis and comment on what you can (or cannot) use its estimates for.
2- if we are skeptical of the current study. Report the power analysis and comment on what you can (or cannot) use its estimates for.
3- if we only have access to 30 participants. Identify the power for each relevant effect and discuss whether it's worth to run the study and why
The list above is also what you should discuss in your code-less report.


## Learning objectives

- Learn how to calculate statistical power
- Critically appraise how to apply frequentist statistical power

### Exercise 1

How much power does your study have (if your model estimates are quite right)?
- Load your dataset (both training and testing), fit your favorite model, assess power for your effects of interest (probably your interactions).

```{r}
df_train <- read.csv("cleanedData.csv")
df_test <- read.csv("cleanedDataTest.csv")

# Adding 66
df_test$ID <- df_test$ID+66
df_test
df <- merge(df_train,df_test, all = T)

sub_df <- df %>% 
  dplyr::select(CHI_MLU, Diagnosis, VISIT, vIQ1,ID) %>% 
  na.omit()

m2 <- lmer(CHI_MLU ~
            Diagnosis * VISIT + vIQ1 + 
            (1 + VISIT | ID), 
          sub_df, REML = F)

```

## EFFECT SIZE
```{r}

summary(m2)
# Interaction estimate: #0.2517469

```

If the effect is real, would we have enough power to expect a positive result? 

```{r}

# What is the current effect size for the interaction? 

fixef(m2)["DiagnosisTD:VISIT"] #0.2517469

power_int <- powerSim((m2), fixed("Diagnosis:VISIT"), nsim = 10) #before it said along = ID in the end

print(power_int) #100%

power_vIQ <- powerSim(m2, fixed("vIQ1"), nsim = 50)

print(power_vIQ) #100%


#while (1 == 1){
#  print("JULIA GET OUT")
#}

```

- Report the power analysis and comment on what you can (or cannot) use its estimates for.

- Test how many participants you would have to have to replicate the findings (assuming the findings are correct)

```{r}
# TEST
pc2 <- powerCurve((m2), fixed("Diagnosis:VISIT"), nsim =50, along = "ID")

print(pc2)

fixef(m2)["DiagnosisTD:VISIT"]

pc2 <- powerCurve(m2,fixed("Diagnosis:VISIT"), nsim = 10)

```


N.B. Remember that main effects are tricky once you have interactions in the model (same for 2-way interactions w 3-way interactions in the model). If you want to test the power of main effects, run a model excluding the interactions.
N.B. Check this paper: https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12504
You will be using:
- powerSim() to calculate power
- powerCurve() to estimate the needed number of participants
- extend() to simulate more participants

```{r}

```


### Exercise 2

How would you perform a more conservative power analysis?
- Identify and justify a minimum effect size for each of your relevant effects
- take the model from exercise 1 and replace the effects with the minimum effect size that you'd accept.
- assess the power curve by Child.ID, identifying an ideal number of participants to estimate each effect
- if your power estimates do not reach an acceptable threshold simulate additional participants and repeat the previous analysis
- Report the power analysis and comment on what you can (or cannot) use its estimates for.

```{r}

# Reder the model first
m3 <- m2

# DECIDE MINIMUM EFFECT SIZE FOR THE EFFECT WE WANT TO INVESTIGATE
fixef(m3)["DiagnosisTD:VISIT"] <- 0.1

# HOW POWERFUL IT IS WITH THE CURRENT DATA?

powerSim_m3 <- powerSim(m3, fixed("Diagnosis:VISIT"), nsim = 100)
print(powerSim_m3)

# HOW MANY PARTICIPANTS WOULD WE NEED TO GET THE POWER WE WANT

pc_m3 <- powerCurve((m3), fixed("Diagnosis:VISIT"), nsim = 10, along = "ID")


```

Trying to see how many participants we would need to detect an effect of 0.05
```{r}
m_0.05 <- m2

fixef(m_0.05)["DiagnosisTD:VISIT"] <- 0.05

# HOW POWERFUL IS IT WITH THE CURRENT DATA? 

powerSim_m_0.05 <- powerSim(m_0.05, fixed("Diagnosis:VISIT"), nsim = 100)

powerSim_m_0.05 #38%

# HOW MANY PARTICIPANTS WOULD WE NEED TO GET 80% POWER? 

powerCurve_m_0.05 <- powerCurve((m_0.05), fixed("Diagnosis:VISIT"), nsim = 10, along = "ID")

powerCurve_m_0.05

# EXTENDING TO 400
m_0.05 <- extend(m_0.05, along = "ID", n = 400)

# HOW MANY PARTICIPANTS WOULD WE NEED TO GET 80% POWER? 
powerCurve_m_0.05_extend <- powerCurve((m_0.05), fixed("Diagnosis:VISIT"), nsim = 50, along = "ID")

powerCurve_m_0.05_extend

plot(powerCurve_m_0.05_extend)

```
want the power estimation to be more conservative - underestimate- we want it not to be exactly 80% 
uncertainty interval

We design the study and estimate higher - 
setting the estimate lower, for it to demand more participants required so that we do not undershoot in terms of sample size



### Exercise 3

Assume you have only the resources to collect 30 kids (15 with ASD and 15 TDs). Identify the power for each relevant effect and discuss whether it's worth to run the study and why

```{r}

sub30 <- merge(
  (sub30TD <- sub_df %>% 
  filter(Diagnosis == "TD", ID < 31)),
  (sub30ASD <- sub_df %>% 
  filter(Diagnosis  == "ASD", ID <31)), all = T)

m_30p <- lmer(CHI_MLU ~ Diagnosis * VISIT + vIQ1 + (1+VISIT|ID), sub30, REML = F)


powerSim_m_30p <- powerSim((m_30p), fixed("Diagnosis:VISIT"), nsim = 100)

powerSim_m_30p

# 
```
