---
title: "SIMR"
author: "Søren Orm Hansen"
date: "9/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(pacman)

p_load(simr, lme4, dplyr)
```

```{r}
df <- read_csv('cleanedData.csv')

sub_df_train <- df %>%
  dplyr::select(c(ID, types_CHI, VISIT, Diagnosis)) %>%
  na.omit()
#mutate(pred_CHI_MLU_m = fitted(m),
#pred_CHI_MLU_gcm = fitted(gcm))

```

```{r}
m1 <-
  lme4::lmer(data = sub_df_train,
             types_CHI ~ VISIT + Diagnosis + (1 + VISIT | ID),
             REML = F)

summary(m1)

powerV <- powerSim(m1, fixed('VISIT'))
powerD <- powerSim(m1, fixed('Diagnosis'))

powerV
powerD
```

```{r}
fixef(m1)['VISIT'] <- 10
fixef(m1)['DiagnosisTD'] <- 30

powerCV <- powerCurve(m1, fixed('VISIT'), along = 'ID', nsim = 50)
powerCD <-
  powerCurve(m1, fixed('Diagnosis'), along = 'ID', nsim = 50)

plot(powerCV)
plot(powerCD)

```

```{r}
m2 <- extend(m1, along = "ID", n = 120)
#pc3 <- powerCurve(m2, fixed('Diagnosis'), along="ID", nsim = 50)
plot(pc3)
```



```{r}

simdata <- simdata

model1 <- glmer(z ~ x + (1|g), family="poisson", data=simdata)

summary(model1)

fixef(model1)["x"] <- -0.05

summary(model1)

powerSim(model1)

```


Power for predictor 'x', (95% confidence interval):====================================================|
      31.30% (28.43, 34.28)

Test: z-test
      Effect size for x is -0.050

Based on 1000 simulations, (4 warnings, 0 errors)
alpha = 0.05, nrow = 30

Time elapsed: 0 h 1 m 22 s

```{r}
model2 <- extend(model1, along="x", n=20)
powerSim(model2)
```

Power for predictor 'x', (95% confidence interval):====================================================|
      94.60% (93.01, 95.92)

Test: z-test
      Effect size for x is -0.050

Based on 1000 simulations, (22 warnings, 0 errors)
alpha = 0.05, nrow = 60

Time elapsed: 0 h 1 m 20 s

```{r}
pc2 <- powerCurve(model2)
print(pc2)
```
Power for predictor 'x', (95% confidence interval),
by largest value of x:
      3:  4.50% ( 3.30,  5.98) - 9 rows
      5:  8.30% ( 6.66, 10.19) - 15 rows
      7: 12.80% (10.79, 15.03) - 21 rows
      9: 25.40% (22.73, 28.22) - 27 rows
     11: 41.30% (38.23, 44.42) - 33 rows
     12: 49.60% (46.46, 52.75) - 36 rows
     14: 66.00% (62.97, 68.94) - 42 rows
     16: 80.20% (77.59, 82.63) - 48 rows
     18: 89.00% (86.89, 90.87) - 54 rows
     20: 95.90% (94.48, 97.04) - 60 rows

Time elapsed: 0 h 9 m 18 s








